Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-01-11 vesicle_code/docs/adhesion/adhesion.tex
worked on extensional example

%
Adhesion between a vesicle and a substrate has been extensively studied, primarily focused on analyzing the static equilibrium shapes \cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin, LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase} than the the transient adhesion process; exceptions include \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by Evans and Metcalfe \cite{EvansMetcalfe1984_BJ}, who measured 
the reduction in the free energy per unit area of membrane-membrane contact formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion energy per unit area to the bending energy (bending rigidity of lipid bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle deformation in the presence of adhesion \cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In this work we focus on regimes where the ratio is of order unity, between the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$) regimes. 
In this regime the membrane deformation may increase the ``contact area'' between two vesicles and enhance the adhesion effects. The equilibration of vesicle membranes under adhesion in these regimes have been well studied and documented (see \cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and references therein). However it is unclear how the adhesion couples to the vesicle hydrodynamics in such regime.  Our work in this paper aims to address this question with quantitative characterization in terms of physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a doublet due to their Van der Waals' attractive interactions \cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. Due to the strong Van der Waals' adhesive force, the vesicles are far from spherical shape and the membrane is almost flat in the ``contact" region. Gires {\it et al.} used small-deformation analysis to investigate the hydrodynamic
interactions between two vesicles in a planar shear flow with a long separation distance \cite{GiresDankerMisbah2012_PRE}. They found that the vesicle interaction could be either attractive or repulsive depending on the organization of the two vesicles relative to the shear flow. To the best of authors' knowledge,
the effects of close-range vesicle adhesion on their hydrodynamics under an external flow have not been studied and quantified. The goal of this work is to use
state-of-the-art boundary integral simulations to numerically investigate the dynamics of two vesicles in both planar shear flow and extensional flow for a wide range of vesicle shapes and adhesive strength and distance.

@@

>>>>

%
Adhesion between a vesicle and a substrate has been extensively studied, primarily focused on analyzing the static equilibrium shapes \cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin, LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase} than the the transient adhesion process; exceptions include \cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by Evans and Metcalfe \cite{EvansMetcalfe1984_BJ}, who measured 
the reduction in the free energy per unit area of membrane-membrane contact formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion energy per unit area to the bending energy (bending rigidity of lipid bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle deformation in the presence of adhesion \cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In this work we focus on regimes where the ratio is of order unity, between the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$) regimes. 
In this regime the membrane deformation may increase the ``contact
area'' between two vesicles and enhance the adhesion effects. The
equilibration of vesicle membranes under adhesion in these regimes have
been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

In a quiescent environment, sub-micron size vesicles are found to form a
doublet due to their Van der Waals' attractive
interactions~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.
Due to the strong Van der Waals' adhesive force, the vesicles are far
from spherical shape and the membrane is almost flat in the ``contact"
region. Gires {\it et al.} used small-deformation analysis to
investigate the hydrodynamic interactions between two vesicles in a
planar shear flow with a long separation distance
\cite{GiresDankerMisbah2012_PRE}. They found that the vesicle
interaction could be either attractive or repulsive depending on the
organization of the two vesicles relative to the shear flow. To the best
of authors' knowledge, the effects of close-range vesicle adhesion on
their hydrodynamics under an external flow have not been studied and
quantified. The goal of this work is to use state-of-the-art boundary
integral simulations to numerically investigate the dynamics of two
vesicles in both planar shear flow and extensional flow for a wide range
of vesicle shapes and adhesive strength and distance.

@@

In Figure~\ref{fig:extensionalInclinationAngle}, we plot the inclination
angle of a doublet formed by two vesicles of reduced area $0.70$ at four
different extensional rates.

\begin{figure}[htp]
  \includegraphics[width=0.45\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \todo[inline]{need a plot describing what the inclination angle
  is}
  \caption{\label{fig:extensionalInclinationAngle} The inclination angle
  of a doublet formed by vesicles of reduced area $0.70$ submerged in an
  extensional flow with an extensional rate of $\chi$.}
\end{figure}
